---
title: "Pain Reduction Analysis"
author: "Rumil Legaspi"
date: "8/26/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```




# Pain Reduction Analysis Project

## Introduction
  
Pain reduction - the primary efficacy endpoint of the QUEST Study. Understanding the level of pain reduced holds utmost importance, not only for the clinical team at NEUROS, but more importantly for the under-served post-amputee population seeking treatment. In nearing the end of the study, we have seen an increase in the passing of subjects within the eligibility phase, specifically, during the saline and lidocaine injections. This short analysis aims to explore the Saline and Lidocaine injections of subjects within the years of 2019 to the month of August 2021, using data extracts derived from the EDC system, Clindex.

# Objectives and Questions
The following are a few questions this analysis seeks to better understand:

- On average, what is the percent in pain reduction subjects are receiving post-saline and lidocaine injection? By Site?

- What do the drops in pain look like? Specifically, how do sites compare and look like.

- Based on the passing were seeing how are they passing and if failing how are they failing?

- Filter by year 2019,2020, and 2021


- What are the pass rates in 2019-2021 by year (the cut from 23rd aug extract)

- how many passed saline, how many passed lidocaines, for all of them
- based on subjects that passed how much on averae did they pass by?
-Who failed and passed? by year, and in total (all years 2019-2021 combined)

- then only subjects who passed, I take 0 minute saline mark in those 3 years and that 0 min    column minus 20 minutes to
    - see what verage pain decraease is, percentage wise
    - max min, medium,, basic stats
    - highest performing and least performing sites



## In short...
- sort who passed
- analyze pain decrease
- are subjects blowing out water? decrease in pain 70,80,90%?
- what distribution looka like?
- How does it look relative to 2019-2020-2021
- clues on our 3% increase in injection rates

#  Loading in Libraries
```{r}
library(tidyverse)
library(lubridate)
library(readr)
```

# Filepaths
```{r}
pc <-
```


# 1. Importing Data
```{r}
# Evaluation Visit
eval_visit <- read_csv("Data/nm_evalvis (5).csv", locale = locale(encoding = "Latin1"))
View(nm_evalvis_5_)

# Saline Injections
saline_inj <- read_csv("Data/nm_saline.csv")
View(saline_inj)

# Lidocaine Injections
lidocaine_inj <- read_csv("Data/nm_lidocain.csv")
View(lidocaine_inj)
```

```{r}
is_tibble(eval_visit)
```


# Joining Datasets
```{r}
# left Joining saline and Lidocaine datasets first

glimpse(saline_inj)

glimpse(lidocaine_inj)

sal_lido_join <- left_join(saline_inj, lidocaine_inj, by = "KEYID") %>% 
  glimpse()

# left joining saline and lido dataset with evaluation visit dataset

inj_data <- left_join(sal_lido_join, eval_visit, by = "KEYID") %>%  
  glimpse()

```



# 2. Data Cleaning
```{r}
# Selecting & renaming Columns
cleaning_inj_data <- inj_data %>% 
  rename(key_id              = KEYID,
         site                = `Site #`,
         subj_id             = `Unique Subject ID`,
         visit_date          = `Date of visit`,
         sal_0min            = `0 min.x`,
         sal_5min            = `+ 5 min.x`,
         sal_10min           = `+ 10 min.x`,
         sal_15min           = `+ 15 min.x`,
         sal_20min           = `+ 20 min.x`,
         lido_0min           = `0 min.y`,
         lido_5min           = `+ 5 min.y`,
         lido_10min          = `+ 10 min.y`,
         lido_15min          = `+ 15 min.y`,
         lido_20min          = `+ 20 min.y`,
         inj_eval_num        = `Injection Evlaution Form Number`,
         meds_today          = `N/A, None today`,
         med_name            = Name,
         med_dose            = Dose,
         med_time            = Time,
         time_intl_pain_lvl  = `Time of Pain rating`,
         intl_pain_lvl       = `What is subject<U+0092>s Overall limb pain right now?`,
         sal_inj_not_done    = `Saline Injection Not Done` ,
         inj_location        = `Target of injection`,
         sal_inj_start_time  = `Time of saline injection`,
         sal_inj_end_time    = `End of Saline Assessment Time`,
         sal_eval_outcome    = `Saline Injection Outcome?`,
         lido_inj_start_time = `Time of lidocaine injection`,
         lido_inj_end_time   = `End of Lidocaine Assessment Time`,
         lido_eval_outcome   = `Lidocaine Injection Outcome` 
         
         ) %>%  glimpse()
  
  # Cleaning data types
cleaning_inj_data <- as.factor(cleaning_inj_data$sal_0min)
glimpse(cleaning_inj_data)
 

typeof(cleaning_inj_data$sal_0min)
#should i go integer or use double?
 sal_5min,
            sal_10min,
            sal_15min,
            sal_20min,
            lido_0min,
            lido_5min,
            lido_10min,
            lido_15min,
            lido_20min,
            intl_pain_lvl))  
  
#is nRS pain factor?
  
```

eval_visit <- eval_visit %>% rename(target_inj = `Target of injection`)

