---
title: "Subject Pain Distribution Histograms"
output: html_notebook
---
#Injection Analyses


## Subject Pain Distribution Histograms


```{r}

prep.Inj_data_plots_14_21 <- cleaned_data_14_21[-c(1:3, 15,22,33:35)]

prep.Inj_data_plots_14_21 <- prep.Inj_data_plots_14_21  %>% 
  mutate('1sal_00min' = sal_00min,
         '1sal_05min' = sal_05min,
         '1sal_10min' = sal_10min,
         '1sal_15min' = sal_15min,
         '1sal_20min' = sal_20min,
         '2lido_00min' = lido_00min,
         '2lido_05min' = lido_05min,
         '2lido_10min' = lido_10min,
         '2lido_15min' = lido_15min,
         '2lido_20min' = lido_20min)
 
  
  Inj_data_plots_14_21 <- prep.Inj_data_plots_14_21 %>% 
    select('1sal_00min',
         '1sal_05min',
         '1sal_10min',
         '1sal_15min',
         '1sal_20min',
         '2lido_00min',
         '2lido_05min',
         '2lido_10min',
         '2lido_15min',
         '2lido_20min')  %>% 
   keep(is.numeric) %>%                       # Using numeric columns
  gather() %>%  
  ggplot(aes(value)) +                       #plotting
    facet_wrap(~ key, scales = "free", ncol = 5) +
    geom_bar(fill = "#F8766D", color = "black")  + 
    labs(title="Injection Pain Distributions", 
         subtitle="JAN2014-AUG2021",
         caption="Source: Lidocaine & Saline Pain Levels 0 to 20 minutes",
         x="Reported Subject Pain Levels During Injections",
         y="Frequency")
                 

#2014-2021
Inj_data_plots_14_21



prep.Inj_data_plots_19_21 <- cleaned_data_19_21[-c(1:3, 15,22,33:35)]


prep.Inj_data_plots_19_21 <- prep.Inj_data_plots_19_21 %>% 
  mutate('1sal_00min' = sal_00min,
         '1sal_05min' = sal_05min,
         '1sal_10min' = sal_10min,
         '1sal_15min' = sal_15min,
         '1sal_20min' = sal_20min,
         '2lido_00min' = lido_00min,
         '2lido_05min' = lido_05min,
         '2lido_10min' = lido_10min,
         '2lido_15min' = lido_15min,
         '2lido_20min' = lido_20min)
                                                                     

  Inj_data_plots_19_21 <- prep.Inj_data_plots_19_21 %>% 
    select('1sal_00min',
         '1sal_05min',
         '1sal_10min',
         '1sal_15min',
         '1sal_20min',
         '2lido_00min',
         '2lido_05min',
         '2lido_10min',
         '2lido_15min',
         '2lido_20min')  %>% 
   keep(is.numeric) %>%                       # Using numeric columns
  gather() %>%
  ggplot(aes(value)) +                       #plotting
    facet_wrap(~ key, scales = "free",nrow = 2, ncol = 5) +
    geom_bar(fill = "#00BFC4", color = "black")  +
    labs(title="Injection Pain Distributions of All Subjects",
         subtitle="JAN2019 - AUG2021 \nHistograms of Subject Reported Pain Levels",
         caption="Source: Lidocaine & Saline Pain Levels 0 to 20 minutes",
         x="Reported Subject Pain Levels During Injections",
         y="Frequency")



#2019-2021
Inj_data_plots_19_21



#2019 Only


prep.Inj_data_plots_19 <- cleaned_data_19[-c(1:3, 15,22,33:35)]

prep.Inj_data_plots_19 <- prep.Inj_data_plots_19 %>% 
  mutate('1sal_00min' = sal_00min,
         '1sal_05min' = sal_05min,
         '1sal_10min' = sal_10min,
         '1sal_15min' = sal_15min,
         '1sal_20min' = sal_20min,
         '2lido_00min' = lido_00min,
         '2lido_05min' = lido_05min,
         '2lido_10min' = lido_10min,
         '2lido_15min' = lido_15min,
         '2lido_20min' = lido_20min)
                                                                     

  Inj_data_plots_19 <- prep.Inj_data_plots_19 %>% 
    select('1sal_00min',
         '1sal_05min',
         '1sal_10min',
         '1sal_15min',
         '1sal_20min',
         '2lido_00min',
         '2lido_05min',
         '2lido_10min',
         '2lido_15min',
         '2lido_20min')  %>% 
   keep(is.numeric) %>%                       # Using numeric columns
  gather() %>%
  ggplot(aes(value)) +                       #plotting
    facet_wrap(~ key, scales = "free",nrow = 2, ncol = 5) +
    geom_bar(fill = "#FFDB6D", color = "black")  +
    labs(title="Injection Pain Distributions",
         subtitle="2019 \nHistograms of Subject Reported Pain Levels",
         caption="Source: Lidocaine & Saline Pain Levels 0 to 20 minutes",
         x="Reported Subject Pain Levels During Injections",
         y="Frequency")

  Inj_data_plots_19
  
#2020 Only


prep.Inj_data_plots_20 <- cleaned_data_20[-c(1:3, 15,22,33:35)]

prep.Inj_data_plots_20 <- prep.Inj_data_plots_20 %>% 
  mutate('1sal_00min' = sal_00min,
         '1sal_05min' = sal_05min,
         '1sal_10min' = sal_10min,
         '1sal_15min' = sal_15min,
         '1sal_20min' = sal_20min,
         '2lido_00min' = lido_00min,
         '2lido_05min' = lido_05min,
         '2lido_10min' = lido_10min,
         '2lido_15min' = lido_15min,
         '2lido_20min' = lido_20min)
                                                                     

  Inj_data_plots_20 <- prep.Inj_data_plots_20 %>% 
    select('1sal_00min',
         '1sal_05min',
         '1sal_10min',
         '1sal_15min',
         '1sal_20min',
         '2lido_00min',
         '2lido_05min',
         '2lido_10min',
         '2lido_15min',
         '2lido_20min')  %>% 
   keep(is.numeric) %>%                       # Using numeric columns
  gather() %>%
  ggplot(aes(value)) +                       #plotting
    facet_wrap(~ key, scales = "free",nrow = 2, ncol = 5) +
    geom_bar(fill = "#FFDB6D", color = "black")  +
    labs(title="Injection Pain Distributions",
         subtitle="2020 \nHistograms of Subject Reported Pain Levels",
         caption="Source: Lidocaine & Saline Pain Levels 0 to 20 minutes",
         x="Reported Subject Pain Levels During Injections",
         y="Frequency")

 
 Inj_data_plots_20 
  
   
#2021 Only


prep.Inj_data_plots_21 <- cleaned_data_21[-c(1:3, 15,22,33:35)]

prep.Inj_data_plots_21 <- prep.Inj_data_plots_21 %>% 
  mutate('1sal_00min' = sal_00min,
         '1sal_05min' = sal_05min,
         '1sal_10min' = sal_10min,
         '1sal_15min' = sal_15min,
         '1sal_20min' = sal_20min,
         '2lido_00min' = lido_00min,
         '2lido_05min' = lido_05min,
         '2lido_10min' = lido_10min,
         '2lido_15min' = lido_15min,
         '2lido_20min' = lido_20min)
                                                                     

  Inj_data_plots_21 <- prep.Inj_data_plots_21 %>% 
    select('1sal_00min',
         '1sal_05min',
         '1sal_10min',
         '1sal_15min',
         '1sal_20min',
         '2lido_00min',
         '2lido_05min',
         '2lido_10min',
         '2lido_15min',
         '2lido_20min')  %>% 
   keep(is.numeric) %>%                       # Using numeric columns
  gather() %>%
  ggplot(aes(value)) +                       #plotting
    facet_wrap(~ key, scales = "free",nrow = 2, ncol = 5) +
    geom_bar(fill = "#FFDB6D", color = "black")  +
    labs(title="Injection Pain Distributions",
         subtitle="2021 \nHistograms of Subject Reported Pain Levels",
         caption="Source: Lidocaine & Saline Pain Levels 0 to 20 minutes",
         x="Reported Subject Pain Levels During Injections",
         y="Frequency")
  
Inj_data_plots_21  
```


### filtered by subjects who passed and failed
```{r}
prep.Inj_data_plots_19_21 <- cleaned_data_19_21[-c(1:3, 15,22,33,35)]
colnames(prep.Inj_data_plots_19_21)


prep.Inj_data_plots_19_21 <- prep.Inj_data_plots_19_21 %>% filter(pass == 1) %>% 
  mutate('1sal_00min' = sal_00min,
         '1sal_05min' = sal_05min,
         '1sal_10min' = sal_10min,
         '1sal_15min' = sal_15min,
         '1sal_20min' = sal_20min,
         '2lido_00min' = lido_00min,
         '2lido_05min' = lido_05min,
         '2lido_10min' = lido_10min,
         '2lido_15min' = lido_15min,
         '2lido_20min' = lido_20min)
                                                                     

  Inj_data_plots_19_21_passed <- prep.Inj_data_plots_19_21 %>% 
    select('1sal_00min',
         '1sal_05min',
         '1sal_10min',
         '1sal_15min',
         '1sal_20min',
         '2lido_00min',
         '2lido_05min',
         '2lido_10min',
         '2lido_15min',
         '2lido_20min')  %>% 
   keep(is.numeric) %>%                       # Using numeric columns
  gather() %>%
  ggplot(aes(value)) +                       #plotting
    facet_wrap(~ key, scales = "free",nrow = 2, ncol = 5) +
    geom_bar(fill = "#00BFC4", color = "black")  +
    labs(title="Injection Pain Distributions of Passed Subjects",
         subtitle="JAN2019 - AUG2021 \nHistograms of Subject Reported Pain Levels",
         caption="Source: Lidocaine & Saline Pain Levels 0 to 20 minutes",
         x="Reported Subject Pain Levels During Injections",
         y="Frequency")
  
  
prep.Inj_data_plots_19_21 <- cleaned_data_19_21[-c(1:3, 15,22,33,35)]
colnames(prep.Inj_data_plots_19_21)


prep.Inj_data_plots_19_21 <- prep.Inj_data_plots_19_21 %>% filter(pass == 0) %>% 
  mutate('1sal_00min' = sal_00min,
         '1sal_05min' = sal_05min,
         '1sal_10min' = sal_10min,
         '1sal_15min' = sal_15min,
         '1sal_20min' = sal_20min,
         '2lido_00min' = lido_00min,
         '2lido_05min' = lido_05min,
         '2lido_10min' = lido_10min,
         '2lido_15min' = lido_15min,
         '2lido_20min' = lido_20min)
                                                                     

  Inj_data_plots_19_21_failed <- prep.Inj_data_plots_19_21 %>% 
    select('1sal_00min',
         '1sal_05min',
         '1sal_10min',
         '1sal_15min',
         '1sal_20min',
         '2lido_00min',
         '2lido_05min',
         '2lido_10min',
         '2lido_15min',
         '2lido_20min')  %>% 
   keep(is.numeric) %>%                       # Using numeric columns
  gather() %>%
  ggplot(aes(value)) +                       #plotting
    facet_wrap(~ key, scales = "free",nrow = 2, ncol = 5) +
    geom_bar(fill = "#00BFC4", color = "black")  +
    labs(title="Injection Pain Distributions of Failed Subjects",
         subtitle="JAN2019 - AUG2021 \nHistograms of Subject Reported Pain Levels",
         caption="Source: Lidocaine & Saline Pain Levels 0 to 20 minutes",
         x="Reported Subject Pain Levels During Injections",
         y="Frequency")  
  
  
  
  
  Inj_data_plots_19_21
  Inj_data_plots_19_21_passed
  Inj_data_plots_19_21_failed
```

# Points
- Notice the pain levels for saline injections for passed subject remain relatively the same in comparison failed subjects whose pain distrbution start to skew towards the right or kind of just go all over the place.


How much of the subjects that passed, passed lidocaine with an even or odd number?

